---
html_document:
  toc: yes
  toc_depth: '3'
  df_print: paged
author: "Munther Alnghemshi"
date: "11/19/2020"
title: "Diamond analysis"
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#packages
library(MASS)
library(ggplot2)
library(dplyr)
library(data.table)
library(tidyr)
library(tidyverse)
library(DT)
library(pander)
```
\
\
\
\
\

### Table
'
```{r, echo=FALSE, warning=FALSE, message=FALSE}
head(diamonds) #show head of data table
```

### Numerical values
```{r, echo=FALSE, warning=FALSE, message=FALSE}
diamonds %>%
  summarise(mean = mean(price), sd = sd(price), median = median(price)) #avg and median of prices 


diamonds %>% #summurize the frequency of each cut 
  group_by(cut) %>%
  summarise(counts = n())

```

### Representation for categorical variable of colors
```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(data = diamonds, aes(x = color)) + #rep of colo
  geom_bar()
```


#### Price vs Number of Diamonds:
```{r, echo=FALSE, warning=FALSE, message=FALSE}
qplot(x = price, data = diamonds, fill = I('blue'), color = I('black'),
      binwidth = 450)+
  scale_x_continuous(breaks = seq(0,20000, 2000))+
  ylab('Number of Diamonds')+
  xlab('Price')
```





Through a box plot, we observe that the highest price diamonds are premium cut, lowest price diamonds are ideal cut.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
diamonds %>%
  ggplot(aes(x=cut,y=price, color=cut)) +
  geom_boxplot() #color vs price
```


***
\

#### Color vs Number of Diamonds using Box-Plot

```{r, echo=FALSE, warning=FALSE, message=FALSE}
qplot(x = color, data = diamonds, y = price, geom = 'boxplot') +
  coord_cartesian(ylim = c(0, 10000))+
  ylab('No. of Diamonds')+
  xlab('Color')
```

***

### The price of a diamond and it’s carat weight


```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(diamonds,aes(x=carat,y=price))+
  geom_point(color='blue',fill='blue')+ #carat vs price
  xlim(0,quantile(diamonds$carat,0.99))+
  ylim(0,quantile(diamonds$price,0.99))+
  ggtitle('Diamond price vs. carat')
```


***
\
\

### Plots of each combination of x, y, and z
\
\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
#scatter plots of each combination of x, y, and z

ggplot(diamonds, aes(x = x, y = y)) +
  geom_point()
ggplot(diamonds, aes(x = x, y = z)) +
  geom_point()
ggplot(diamonds, aes(x = y, y = z)) +
  geom_point()
```

### Price per Carat with respect to the cut


```{r, echo=FALSE, warning=FALSE, message=FALSE}
qplot(x = cut, y = price/carat, data = diamonds, geom = 'boxplot')+
  ylab('Price/Carat')+
  xlab('cut')
```



#### Cut distribitions 


```{r, echo=FALSE, warning=FALSE, message=FALSE}
qplot(x = price, data = diamonds, fill = I('blue'), color = I('black')) +
  facet_wrap(~cut, ncol = 2)+
  ylab('Number of Diamonds')+
  xlab('Price')
```



## Inferential Statistics


### The one-way ANOVA summary:


```{r, echo=FALSE, warning=FALSE, message=FALSE}
 
dia.aov <- aov(carat ~ cut, data=diamonds)
summary(dia.aov) %>% pander()
 

```

***

###Tukey’s Post-hoc test: 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
tu <- TukeyHSD(dia.aov)
datatable(tu$cut)
```

